---
title: "nf_xpatial | MapMyCells"
output:
  html_document:
    code_folding: hide
    css: "style.css"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview - MapMyCells with output from `nf_xpatial`

* Illustration purposes of using nf_xpatial outputs for further processing including cell label transfer.
* This example uses the object which was area normalized, highlighting that mapmycells require the raw data (which is kept in the object under the original `Xenium` assay in the `counts` slot)
* Information about references can be found at: <https://portal.brain-map.org/atlases-and-data/bkp/mapmycells/cell-type-references>


```{r load libraries, message=FALSE}
set.seed(1234)
library(Seurat)
library(ggplot2)
library(anndata)
```

```{r}
inputdir <- "./data/"
outdir <- "./results/1_MapMyCells/"
dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
```


```{r}
seurat_obj <- readRDS(file = paste0(inputdir,"compiled_area_norm_all_clusters.rds"))

DefaultAssay(object = seurat_obj) <- "Xenium"

# for AreaNorm only! Need to JoinLayers for Xenium assay (this is automatically done for LogNorm).
seurat_obj[["Xenium"]] <- JoinLayers(seurat_obj[["Xenium"]])

seurat_obj
```

# Save metadata for raw_data_checks (a sanity check):

```{r}
meta_area <- as.data.frame(seurat_obj@meta.data)

saveRDS(meta_area, paste0(outdir,"meta_area.rds"))
```

# Save raw counts to `anndata` format:

```{r}
data_matrix_t_area <- Matrix::t(seurat_obj[["Xenium"]]$counts)

#raw matrix across all samples
saveRDS(data_matrix_t_area, paste0(outdir,"data_matrix_t_area.rds"))
```

```{r}
# Convert to anndata format
ad_area <- AnnData(
  X = data_matrix_t_area,
  obs = data.frame(group = rownames(data_matrix_t_area), row.names = rownames(data_matrix_t_area)),
  var = data.frame(type = colnames(data_matrix_t_area), row.names = colnames(data_matrix_t_area))
)

# Write to compressed h5ad file
write_h5ad(ad_area, paste0(outdir,"seurat_integrated_mapmycells_area.h5ad"),compression='gzip')

# Check file size. File MUST be <2GB to upload for MapMyCells
print(paste("Size in MB:", round(file.size(paste0(outdir,"seurat_integrated_mapmycells_area.h5ad"))/2^20)))
```